Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Chinook Salmon Populations

### 2022-02-01 analysis/paper/paper.Rmd
cleaned up some references

# sum fall and spring impacts
total_impacts = impact_df %>%
  group_by(perc_diet, p_chnk) %>%
  summarise(across(c(ends_with("_N"),
                     ends_with("_Lci"),
                     ends_with("_Uci")),
                   sum),
            .groups = "drop")

@@

During the fall DSR emigration, we estimated an average-size Northern Pikeminnow (***value*** mm) to consume ***value*** g of fish to maintain their body size, assuming 60% of their diet consists of fish prey (Figure \@ref(fig:bio-p)). Fall consumption ranged from ***value*** g for a diet of 30% fish to ***value*** g for a diet of 90% fish. During the spring NRR emigration, we estimated an average-size Northern Pikeminnow (***value***) to consume ***value*** 44.29 g of fish, assuming 60% of their diet is fish prey (Figure \@ref(fig:bio-p)). Spring consumption ranged from ***value*** g for a diet of 30% fish to ***value*** g for a diet of 90% fish. Given estimated mean Northern Pikeminnow population sizes of ***value*** from fall estimates and ***value*** for spring 2021, and assuming 60% of their diet is fish, we estimate that Northern Pikeminnow can consume ***value*** metric tons of fish during the fall DSR emigration and an additional ***value*** metric tons during the spring NRR emigration in Deadwater Slough. In total, the Northern Pikeminnow population can consume ***value*** metric tons of fish during the DSR and NRR emigrations to maintain their population size, assuming 60% of their diet consists of fish.

<!--
We estimated an individual, average-size (`r avg_fall_NPM_TL` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Fall"])` g) of fish during the DSR emigration to maintain their body size, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)). Similarly, we estimated an average-size (`r avg_spring_NPM_TL` mm) Northern Pikeminnow to consume `r median(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g (range: `r min(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` - `r max(bio_g_summ$cumm_g[bio_g_summ$season == "Spring"])` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)). Given a mean Northern Pikeminnow population size of `r prettyNum(unique(impact_df$N[impact_df$season == "Fall"]), big.mark = ",")` from fall estimates and `r prettyNum(unique(impact_df$N[impact_df$season == "Spring"]), big.mark = ",")` for spring 2021 and using the median value from diet scenarios (i.e., 50% of their diet consists of fish), we'd estimate that the Northern Pikeminnow population can consume `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 50]), big.mark = ",")` g (range: `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 10]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 10]), big.mark = ",")` - `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 90]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 90]), big.mark = ",")` g) fish during the DSR emigration and an additional `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 50]), big.mark = ",")` g (range: `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 10]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 10]), big.mark = ",")` - `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 90]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 90]), big.mark = ",")` g) during the NRR emigration in Deadwater Slough. In total, the Northern Pikeminnow population can consume `r prettyNum(unique(impact_df$g_fish_cons[impact_df$season == "Fall" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Fall" & impact_df$perc_diet == 50]) + unique(impact_df$g_fish_cons[impact_df$season == "Spring" & impact_df$perc_diet == 50]) * unique(impact_df$N[impact_df$season == "Spring" & impact_df$perc_diet == 50]), big.mark = ",")` g of fish during the DSR and NRR emigrations to maintain their population size if we assume 50% of their diet consists of fish.
-->

@@

Using the total g of fish consumed by the Northern Pikeminnow population (summing fall and spring surveys) we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmn make up 30%, 50%, and 65% of total fish prey consumed (Figure ***reference***). Assuming that 50% of the total fish consumed are juvenile Chinook Salmon, we estimate between (30% fish scenario with CIs) and (90% fish scenario with CIs) juvenile Chinook Salmon will be consumed. We then used our estimates of total juvenile Chinook Salmon consumed to estimate "adult equivalents" using the Granite-to-Granite smolt-to-adult survival rate of ***value*** (SE: ***value***; Table ***reference***). Assuming predation in Deadwater Slough was reduced or eliminated, the juvenile consumption would be equivalent to ***value*** adults in the spring (range ***value*** - ***value***) and ***value*** adults (range ***value*** - ***value***) in the fall, totaling ***value*** adult Chinook Salmon.

>>>>

# sum fall and spring impacts
total_impacts = impact_df %>%
  group_by(perc_diet, p_chnk) %>%
  summarise(across(c(ends_with("_N"),
                     ends_with("_Lci"),
                     ends_with("_Uci")),
                   sum),
            .groups = "drop") %>%
  select(perc_diet,
         p_chnk,
         juv_N,
         juv_Lci,
         juv_Uci,
         adult_N,
         adult_Lci,
         adult_Uci)

@@

```{r tons-consumption}
tot_g_DSR = (bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"] * mean(all_est$N[all_est$model == "Schnabel"])) / 1000000
tot_g_NRR = (bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f60_g"] * round(spring_abund$N_S)) / 1000000

During the fall DSR emigration, we estimate an average-size Northern Pikeminnow (`r round(avg_fall_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"]` g of fish to maintain their body size, assuming 60% of their diet consists of fish prey (Figure \@ref(fig:bio-p)). Fall consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f30_g"]` g for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f90_g"]` g for a diet of 90% fish. During the spring NRR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_spring_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f60_g"]` g of fish, assuming 60% of their diet is fish prey (Figure \@ref(fig:bio-p)). Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish. Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` from fall estimates and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` for spring 2021, and assuming 60% of their diet is fish, we estimate that Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and an additional `r round(tot_g_NRR, 1)` metric tons during the NRR emigration in Deadwater Slough.

@@

Using the total fish consumed by the Northern Pikeminnow population (summing fall and spring surveys), we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmon make up 30%, 50%, and 65% of total fish prey consumed (Figure \@ref(fig:impact-p)). Assuming that 50% of the total fish consumed are juvenile Chinook salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed. We then used our estimate of total juvenile Chinook Salmon consumed to estimate an "adult equivalents" using the median Granite-to-Granite SAR of `r round(sar$medsar, 5)` (sd = `r format(round(sar$sdsar, 5), scientific = F)`) from @McCann2019. Assuming predation was non-existent in Deadwater Slough, the juvenile consumption would be equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$p_chnk == 0.50]), big.mark = ",")`).
